\section{Contributions and acknowledgements}
\pg{Relative contributions}
IR conceptualised the project, proposed the coupling mechanisms in Defs \ref{def:coupled_norms_def} and \ref{def:coupled_grfs_def}, proved the major theoretical contributions, ran the GRF experiments and wrote the manuscript. 
SM designed and ran the RFF and RLF GP experiments (Sec.~\ref{sec:rff_rlf_exps}), helped shape the project's direction and made core contributions to the text. 
KC acted as the senior lead, providing technical guidance and developing the algorithms for the matching problem in App.~\ref{sec:quad_grfs}.
RET met frequently throughout the project, giving important advice and support. 
AW provided helpful discussion, supervision and feedback on the manuscript.  

\pg{Acknowledgements and funding}
IR acknowledges support from a Trinity College External Studentship. 
Part of the work was completed as a student researcher at Google.
SM acknowledges funding from the Vice Chancellor’s and the George and Marie Vergottis scholarship of the Cambridge Trust, and the Qualcomm Innovation Fellowship.
RET is supported by Google, Amazon, ARM, Improbable and an EPSRC grant EP/T005386/1.
AW acknowledges support from a Turing AI fellowship under grant EP/V025279/1 and the Leverhulme Trust via CFI.

We thank Bruno Mlodozeniec for his suggestion to use copulas in Sec.~\ref{sec:copulas} and Mark Rowland for insightful discussions about multi-marginal optimal transport and the limitations of pointwise variance reduction. 
Viacheslav Borovitskiy helped guide our discussion of scalable graph-based GPs and, together with Iskander Azangulov, kindly provided updated code for loading the traffic data graph in Sec.~\ref{sec:graph_gp_experiments} and App.~\ref{sec:traffic_expt}.
We thank Matt Ashman and Arijit Sehanobish for their thoughtful feedback on the text, and Jihao Andreas Lin for interesting suggestions about possible GP applications.